---
title: "dplyr"
output: html_document
---

Load the dplyr package.

```{r}
library("dplyr")
# install.packages("dplyr")
```

Load our altmetrics data.

```{r}
counts_raw <- read.delim("data/counts-raw.txt.gz")
```

Only investigate articles that are primary reserarch articles.

```{r}
research <- filter(counts_raw, articleType == "Research Article")
dim(research)
```

What about hte year 2006?

```{r}
research_2006 <- filter(research, year == 2006)
dim(research_2006)
```

Articles in 2006 that have a facebook commment or a tweet.

```{r}
research_2006_fb_tweet <- filter(research, year == 2006,
                                 backtweetsCount > 0 |
                                 facebookCommentCount > 0)
dim(research_2006_fb_tweet)
```

How to select columns?

Obtain article information:

```{r}
article_info <- select(counts_raw, doi, pubDate, journal,
                       title, articleType, authorsCount)
colnames(article_info)
article_info <- select(counts_raw, doi:authorsCount)
colnames(article_info)
```

Obtain metrics:

```{r}
metrics <- select(counts_raw, contains("Count"),
                  -authorsCount, -contains("facebook"))
colnames(metrics)
dim(metrics)
```

In the terminal, the pipe command is |
In dplyr, the pipe commmad is %>%

Obtain articles from 2006 but only the facebook data.

```{r}
facebook_2006 <- research %>% filter(year == 2006) %>%
  select(contains("facebook"))
dim(facebook_2006)
colnames(facebook_2006)
```

Confirming results as you build a dplyr chain of commands:

```{r}
research %>% filter(year == 2006) %>%
  select(contains("facebook")) %>% head
```

How to sort the data frame? Use the function arrange.

```{r}
research %>% arrange(desc(authorsCount)) %>%
  select(authorsCount) %>% head
# No more NA's after filtering for reserach articles
anyNA(research$authorsCount)
```

```{r}
research[1:10, 1:5]
research %>% slice(1:10) %>% select(1:5)
```

Challenges:

Using a chain of pipes, output the titles of the three research articles with the largest 2011 citation count (wosCountThru2011).

```{r}
research %>% arrange(desc(wosCountThru2011)) %>%
  slice(1:3) %>% select(title)
```

Using a chain of pipes, output the author count (authorsCount), title, journal, and subject tags (plosSubjectTags) of the three research articles with the largest number of authors.

```{r}
research %>% arrange(desc(authorsCount)) %>%
  slice(1:3) %>%
  select(authorsCount, title, journal, plosSubjectTags)
```

How to create new columns? Use the funciton mutate.

```{r}
research <- research %>%
  mutate(weeksSincePublished = daysSincePublished / 7,
         yearsSincePublished = weeksSincePublished / 52)
```

Summarize the data with summarize.

```{r}
research %>% summarize(plos_mean = mean(plosCommentCount),
                       plos_sd = sd(plosCommentCount),
                       num = n(),
                       plos_sem = plos_sd / sqrt(num))
```


```{r}
research %>% 
  group_by(journal, year) %>%
  summarize(plos_mean = mean(plosCommentCount),
            plos_sd = sd(plosCommentCount),
            num = n(),
            plos_sem = plos_sd / sqrt(num))
```











